SSA based global code motion of identical computations

### 2017-04-17 acmart-pacmpl-template.tex
iskill

\GCC{} recently got code-hoisting \cite{GCCCodeHoisting} which is implemented as
part of \GVN{}-\PRE{}: it uses the set of ANTIC\_IN and AVAIL\_OUT value
expressions computed for \PRE{}. ANTIC\_IN[B] contains very busy expressions at
basic block B i.e., values computed on all paths from B to exit and
AVAIL\_OUT[B] contains values which are already available. The algorithm hoists
top down to a predecessor.  It uses ANTIC\_IN[B] to know what expressions will
be computed on every path from B to exit, and can be computed in B.  It uses
AVAIL\_OUT[B] to subtract out those values already being computed.  The cost
function is: for each hoist candidate, if all successors of B are dominated by
B, then we know insertion into B will eliminate all the remaining computations.
It then checks to see if at least one successor of B has the value available.
This avoids hoisting it way up the chain to ANTIC.  It also checks to ensure
that B has multiple successors, since hoisting in a straight line is pointless.
The algorithm continues top down the dominator tree, iterating with \PRE{} until
no more changes.  One advantage of \GCC{} implementation is that it works in
sync with the \GVN{}-\PRE{} such that when new hoisting opportunities are
created by \GVN{}-\PRE{}, code-hoisting will hoist them.

@@

Barany \cite{barany2013} presented a global scheduler with integer linear
programming (ILP) formulation with a goal to minimize register pressure. The
results they got were not very promising. It may be because they only used the
scheduler for smaller functions ($<1000$ instructions) also they compiled the
benchmarks for ARM-Cortex which is more resilient to register pressure because
it has more registers compared to say X86.

@@

The original algorithm presented in \cite{das2012} has mismatched types in terms
of uses and nodes because each node can have many instructions and hence many
uses. Also, while iterating on the dominator of each user of a variable we may
reach to the beginning of a function, in that case the inner while loop needs to
terminate. These two cases were missing from the algorithm and we came across
them during implementation.

>>>>

\GCC{}ecently got code-hoisting \cite{GCCCodeHoisting} which is implemented as
part of \GVN{}-\PRE{}: it uses the set of ANTIC\_IN and AVAIL\_OUT value
expressions computed for \PRE{}. ANTIC\_IN[B] contains very busy expressions at
basic block B i.e., values computed on all paths from B to exit and
AVAIL\_OUT[B] contains values which are already available. The algorithm hoists
top down to a predecessor.  It uses ANTIC\_IN[B] to know what expressions will
be computed on every path from B to exit, and can be computed in B.  It uses
AVAIL\_OUT[B] to subtract out those values already being computed.  The cost
function is: for each hoist candidate, if all successors of B are dominated by
B, then we know insertion into B will eliminate all the remaining computations.
It then checks to see if at least one successor of B has the value available.
This avoids hoisting it way up the chain to ANTIC\_IN[B].  It also checks to
ensure that B has multiple successors, since hoisting in a straight line is
pointless.  The algorithm continues top down the dominator tree, working in
tandem with \PRE{} until no more hoisting is possible.  One advantage of \GCC{}
implementation is that it works in sync with the \GVN{}-\PRE{} such that when
new hoisting opportunities are created by \GVN{}-\PRE{}, code-hoisting will
hoist them.

@@

Barany \cite{barany2013} presented a global scheduler with integer linear
programming (ILP) formulation with a goal to minimize register pressure. The
results they got were not very promising. It may be because they only used the
scheduler for smaller functions ($<1000$ instructions); also, they compiled the
benchmarks for ARM-Cortex which is more resilient to register pressure because
it has more registers compared to X86, for example.

@@

\begin{verbatim}
// return true if op is the last use of the
// operand in I.
bool isKill(Operand op, Instruction I) {
  if (isLiveOutUsingMergeSet(op))
    return false;
  B = basic_block(I);
  for each I1 after I in B:
    if (I1 uses op)
      return false;

return true;
}
\end{verbatim}

The original algorithm presented in \cite{das2012} has mismatched types in terms
of uses and nodes because each node can have many instructions and hence many
uses. Also, while iterating on the dominator of each user of a variable we may
reach to the beginning of a function, in that case the inner while loop needs to
terminate. These two cases were missing from the algorithm and we came across
them during implementation. If a variable is not live-out of a basic block, it
still may be used later in the basic block. To establish whether a use of a
variable is a kill we iterate on all the subsequent instructions in a basic
block checking for uses should isLiveOutUsingMergeSet return false.
